HTW E2 Testing

### 2023-10-09 Analysis/e2_test.qmd
update E2 and E3 results, and re-render

@fig-design-e2 illustrates the design of Experiment 2. The stages of the experiment (i.e. training, testing no-feedback, test with feedback), are identical to that of Experiment 1. The only change is that Experiment 2 participants train, and then test, on bands in the reverse order of Experiment 1 (i.e. training on the softer bands; and testing on the harder bands). 
```{dot}
//| label: fig-design-e2
//| fig-cap: "Experiment 2 Design. Constant and Varied participants complete different training conditions. The training and testing bands are the reverse of Experiment 1. "
//| fig-width: 8.0
//| fig-height: 2.5
//| fig-responsive: false
digraph {
  graph [layout = dot, rankdir = LR]

@@

```{r}
#| label: tbl-e2-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap: ["Constant Testing - Deviation", "Varied Testing - Deviation"]
#| layout-ncol: 2
#| echo: fenced

result <- test_summary_table(testE2, "dist", mfun = list(mean = mean, median = median, sd = sd))
result$constant
result$varied

@@

modelName <- "e2_testDistBand_RF_5K"
e2_distBMM <- brm(dist ~ condit * bandInt + (1 + bandInt|id),
                      data=testE2,file=paste0(here::here("data/model_cache",modelName)),
                      iter=5000,chains=4)
mp2 <- GetModelStats(e2_distBMM) |> kable(escape=F,booktabs=T)
mp2
```

As shown in @tbl-e2-bmm-dist, the posterior distribution for the effect of training condition in Experiment 2 is modestly suggestive of a beneficial effect of training variation (β = -70.0, 95% CI [-156.87, 16.66]), with 94% of the posterior falling in the direction of a varied advantage. (SHOULD PROBABLY DO ALTERNATE ANALYSIS THAT ONLY CONSIDERS THE NOVEL EXTRAPOLATION BANDS)

@@

e2_distBMM |> emmeans( ~condit + bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |>
  ggplot(aes(x = bandInt, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
    ylab("Predicted Deviation") + xlab("Velocity Band")+
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE2$vb), 
                     limits = c(0, 1400)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) 
```

```{r}
#| label: fig-e2-ame
#| fig-cap: E2. Predicted Means Per Condition and Band, and Average Marginal Effect (Constant - Varied)
#| fig-width: 10
#| fig-height: 10
new_data_grid=map_dfr(1, ~data.frame(unique(testE2[,c("id","condit","bandInt")])))

cSamp <- e2_distBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> gather_emmeans_draws()  |>
  group_by(contrast, .draw,bandInt) |> summarise(value=mean(.value), n=n())

 ameBand <- cSamp |> ggplot(aes(x=value,y="")) + 
  stat_halfeye() + 
  geom_vline(xintercept=0,alpha=.4)+
  facet_wrap(~bandInt,ncol=1) + labs(x="Marginal Effect (Constant - Varied)", y= NULL)+
  ggtitle("Average Marginal Effect")

bothConditGM <- e2_distBMM %>%
  epred_draws(newdata = new_data_grid,ndraws = 2000, re_formula = NA) |>
  ggplot(aes(x=.epred,y="Mean",fill=condit)) + 
  stat_halfeye() +facet_wrap(~bandInt, ncol = 1) +
  labs(x="Predicted Deviation", y=NULL)+
  ggtitle("Grand Means") +theme(legend.position = "bottom")

(bothConditGM | ameBand) + plot_layout(widths=c(2,1.0))
```

@@

result <- test_summary_table(testE2, "vx", mfun = list(mean = mean, median = median, sd = sd))
result$constant 
result$varied

@@

```{r}
#| label: tbl-e2-bmm-vx
#| tbl-cap: "Experiment 2. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
e2_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=testE2,file=paste0(here::here("data/model_cache", "e2_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
mt3 <-GetModelStats(e2_vxBMM ) |> kable(escape=F,booktabs=T)
mt3
```

See @tbl-e2-bmm-vx for the full model results. The estimated coefficient for training condition (β = 63.49, 95% CI [-133.97, 113.98]) suggests that the varied group tends to produce harder throws than the constant group. More relevant to the issue of discrimination is the slope on Velocity Band (β = 0.71, 95% CI [.58, .84]), which is quite similar to best fitting slope value obtained in Experiment 1 (β = 0.69)

@@

```{r}
#| label: fig-e2-slope-dist
#| fig-cap: Subset of Varied and Constant Participants with the largest estimated slope values. Red lines represent the best fitting line for each participant, gray lines are 100 random samples from the posterior distribution. Colored points and intervals at each band represent the empirical median and 95% HDI. 
#| fig-height: 9
#| fig-width: 10

@@

```{r}
#| label: fig-e2-indv-slopes
#| fig-cap: Slope estimates by participant - ordered from lowest to highest within each condition. 
#| fig-height: 11

nTop <- 2

combined_df <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt) |>
  right_join(new_data_grid, by = join_by("id"),relationship="many-to-many") |> 
  mutate(
    fixed_effects = b_Intercept +
      (bandInt * b_bandInt) +
      condit_dummy * b_conditVaried + 
      (bandInt * condit_dummy) * `b_conditVaried:bandInt` ,  # Note: Replace : with correct operator
    random_effects = Intercept + bandInt_RF,  # Assuming random effects for intercept and bandInt
    estimate = fixed_effects + random_effects
  )

combined_df |> 
  filter(id %in% (indv_coefs |> filter(rank<=nTop) |> pull(id)) )  |> 
  group_by(id, bandInt) |>
  sample_n(200) |>
  ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50", alpha=.1) +
  geom_abline(data=indv_coefs |> 
                filter(id %in% (indv_coefs |> filter(rank<=nTop) |> pull(id))),aes(intercept=Estimate.Intercept,slope=Estimate.bandInt),color="red") +
  stat_pointinterval(data=testE2 |> filter(id %in% (indv_coefs |> filter(rank<=nTop) |> pull(id))), aes(x=bandInt,y=vx, color=vb)) +
  facet_nested_wrap(vars(condit,id),ncol = 2) +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE2$vb), 
                     limits = c(0, 1400)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +  ylab("X Velocity") + xlab("Band")

@@

```{r}
#| label: fig-e2-indv-slopes-small
#| fig-cap: Subset of Varied and Constant Participants with the smallest estimated slope values. Red lines represent the best fitting line for each participant, gray lines are 200 random samples from the posterior distribution. Colored points and intervals at each band represent the empirical median and 95% HDI. 
#| fig-height: 11

nTop <- 2

combined_df |> 
  filter(id %in% (indv_coefs |> filter(rank>nCond-nTop) |> pull(id)) )  |> 
  group_by(id, bandInt) |>
  sample_n(200) |>
  ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50", alpha=.1) +
  geom_abline(data=indv_coefs |> 
                filter(id %in% (indv_coefs |> filter(rank>nCond-nTop) |> pull(id))),aes(intercept=Estimate.Intercept,slope=Estimate.bandInt),color="red") +
  stat_pointinterval(data=testE2 |> filter(id %in% (indv_coefs |> filter(rank>nCond-nTop) |> pull(id))), aes(x=bandInt,y=vx, color=vb)) +
  facet_nested_wrap(vars(condit,id),ncol = 2) +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE2$vb), 
                     limits = c(0, 1400)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +  ylab("X Velocity") + xlab("Band")

@@

>>>>

@fig-design-e2 illustrates the design of Experiment 2. The stages of the experiment (i.e. training, testing no-feedback, test with feedback), are identical to that of Experiment 1. The only change is that Experiment 2 participants train, and then test, on bands in the reverse order of Experiment 1 (i.e. training on the softer bands; and testing on the harder bands). 
```{dot}
//| label: fig-design-e2
//| fig-cap: "Experiment 2 Design. Constant and Varied participants complete different training conditions. The training and testing bands are the reverse of Experiment 1. "
//| fig-width: 8.0
//| fig-height: 2.5
//| fig-responsive: false
//| column: screen-inset-right
digraph {
  graph [layout = dot, rankdir = LR]

@@

```{r}
#| label: tbl-e2-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap: ["Constant Testing - Deviation", "Varied Testing - Deviation"]
#| echo: fenced

result <- test_summary_table(testE2, "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
result$constant 
result$varied 
# make kable table with smaller font size
# result$constant |> kbl(caption="Constant Testing - Deviation",booktabs=T,escape=F) |> kable_styling(font_size = 7)
```

@@

modelName <- "e2_testDistBand_RF_5K"
e2_distBMM <- brm(dist ~ condit * bandInt + (1 + bandInt|id),
                      data=testE2,file=paste0(here::here("data/model_cache",modelName)),
                      iter=5000,chains=4)
mp2 <- GetModelStats(e2_distBMM) |> kable(escape=F,booktabs=T)
mp2

e2_distBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> gather_emmeans_draws()  |> 
   summarize(median_qi(.value),pd=sum(.value>0)/n()) |>
   select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
   mutate(across(where(is.numeric), \(x) round(x, 2)),
          pd=ifelse(value<0,1-pd,pd)) |>
   kbl(caption="Contrasts")

coef_details <- broom.mixed::tidy(e2_distBMM) |>
  filter(term == "conditVaried") |> 
  select(estimate, conf.low, conf.high) |> mutate(across(where(is.numeric), \(x) round(x, 2)))

As shown in @tbl-e2-bmm-dist, the posterior distribution for the effect of training condition in Experiment 2 is modestly suggestive of a beneficial effect of training variation (β = `r coef_details$estimate`, 95% CI \[`r coef_details$conf.low`, `r coef_details$conf.high`\]), with 94% of the posterior falling in the direction of a varied advantage. (SHOULD PROBABLY DO ALTERNATE ANALYSIS THAT ONLY CONSIDERS THE NOVEL EXTRAPOLATION BANDS)

@@

e2_distBMM |> emmeans( ~condit + bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |>
 condEffects()+
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE2$vb), 
                     limits = c(0, 1400)) 
```

@@

result <- test_summary_table(testE2, "vx","X Velocity" ,mfun = list(mean = mean, median = median, sd = sd))
result$constant 
result$varied

@@

```{r}
#| label: tbl-e2-bmm-vx
#| tbl-cap: "Experiment 2. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
e2_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=testE2,file=paste0(here::here("data/model_cache", "e2_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
mt3 <-GetModelStats(e2_vxBMM ) |> kable(escape=F,booktabs=T)
mt3

cd1 <- get_coef_details(e2_vxBMM, "conditVaried")
sc1 <- get_coef_details(e2_vxBMM, "bandInt")
intCoef1 <- get_coef_details(e2_vxBMM, "conditVaried:bandInt")
```

See @tbl-e2-bmm-vx for the full model results. The estimated coefficient for training condition (β = `r cd1$estimate`, 95% CrI \[`r cd1$conf.low`, `r cd1$conf.high`\]) suggests that the varied group tends to produce harder throws than the constant group. More relevant to the issue of discrimination is the slope on Velocity Band (β = `r sc1$estimate`, 95% CrI \[`r sc1$conf.low`, `r sc1$conf.high`\]), which is quite similar to best fitting slope value obtained in Experiment 1 (β = 0.71).(β = `r intCoef1$estimate`, 95% CrI \[`r intCoef1$conf.low`, `r intCoef1$conf.high`\])

@@

```{r}
#| label: fig-e2-slope-dist
#| fig-cap: Subset of Varied and Constant Participants with the largest estimated slope values. Red lines represent the best fitting line for each participant, gray lines are 100 random samples from the posterior distribution. Colored points and intervals at each band represent the empirical median and 95% HDI. 
#| fig-height: 9
#| fig-width: 10
#| eval: false

@@

```{r}
#| label: tbl-e2-slope-quartile
#| tbl-cap: "Slope coefficients by quartile, per condition"

indv_coefs <- as_tibble(coef(e2_vxBMM)$id, rownames="id")|> 
  select(id, starts_with("Est")) |>
  left_join(e2Sbjs, by=join_by(id) )

random_effects <- e2_vxBMM |> 
  gather_draws(`^r_id.*$`, regex = TRUE, ndraws = 1500) |> 
  separate(.variable, into = c("effect", "id", "term"), sep = "\\[|,|\\]") |> 
  mutate(id = factor(id,levels=levels(testE2$id))) |> 
  pivot_wider(names_from = term, values_from = .value) |> arrange(id,.chain,.draw,.iteration)

 indvDraws <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt,RF_Intercept=Intercept) |>
  right_join(new_data_grid, by = join_by("id")) |> 
  mutate(
    Slope = bandInt_RF+b_bandInt,
    Intercept= RF_Intercept + b_Intercept,
    estimate = (b_Intercept + RF_Intercept) + (bandInt*(b_bandInt+bandInt_RF)) + (bandInt * condit_dummy) * `b_conditVaried:bandInt`,
    SlopeInt = Slope + (`b_conditVaried:bandInt`*condit_dummy)
  )

  indvSlopes <- indvDraws |> group_by(id) |> median_qi(Slope,SlopeInt, Intercept,b_Intercept,b_bandInt) |>
  left_join(e2Sbjs, by=join_by(id)) |> group_by(condit) |>
    select(id,condit,Intercept,b_Intercept,starts_with("Slope"),b_bandInt, n) |>
  mutate(rankSlope=rank(Slope)) |> arrange(rankSlope)   |> ungroup()

  indvSlopes |> mutate(Condition=condit) |>  group_by(Condition) |> 
    reframe(enframe(quantile(SlopeInt, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "SlopeInt")) |> 
  pivot_wider(names_from=quantile,values_from=SlopeInt,names_prefix="Q_") |>
  group_by(Condition) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

@@

@fig-e2-bmm-bx2 shows the distributions of estimated slopes relating velocity band to x velocity for each participant, ordered from lowest to highest within condition. Slope values are lower overall for varied training compared to constant training. Figure Xb plots the density of these slopes for each condition. The distribution for varied training has more mass at lower values than the constant training distribution. Both figures illustrate the model's estimate that varied training resulted in less discrimination between velocity bands, evidenced by lower slopes on average.

```{r}
#| label: fig-e2-bmm-bx2
#| fig-cap: Slope distributions between condition
#| fig-subcap: ["Slope estimates by participant - ordered from lowest to highest within each condition. ", "Destiny of slope coefficients by training group"]
#| fig-height: 11
#| fig-width: 12

  indvSlopes |> ggplot(aes(y=rankSlope, x=SlopeInt,fill=condit,color=condit)) + 
  geom_pointrange(aes(xmin=SlopeInt.lower , xmax=SlopeInt.upper)) + 
  labs(x="Estimated Slope", y="Participant")  + facet_wrap(~condit) +
   ggplot(indvSlopes, aes(x = SlopeInt, color = condit)) + 
  geom_density() + labs(x="Slope Coefficient",y="Density")

@@

```{r}
#| label: fig-e2-indv-slopes
#| fig-cap: Subset of Varied and Constant Participants with the smallest and largest estimated slope values. Red lines represent the best fitting line for each participant, gray lines are 200 random samples from the posterior distribution. Colored points and intervals at each band represent the empirical median and 95% HDI. 
#| fig-subcap: ["subset with largest slopes", "subset with smallest slopes"]
#| fig-height: 11
#| fig-width: 12
nSbj <- 3
indvDraws  |> indv_model_plot(indvSlopes, testE2Avg, SlopeInt,rank_variable=Slope,n_sbj=nSbj,"max")
indvDraws |> indv_model_plot(indvSlopes, testE2Avg,SlopeInt, rank_variable=Slope,n_sbj=nSbj,"min")
